---
title: 'Section 2: Fitting PQI Models'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)

hrr <- read_csv("data/hrr_all_years.csv")
hrr$expansion_status <- factor(hrr$expansion_status)
# Reduce to 2018 and evaluate:
hrr <- hrr %>% filter(year == 2018)
```

### PQIs of Interest  

For these models, we will be looking at modeling the relationship between various PQIs at the HRR (hospital-referral region) level and Medicaid expansion status. For model selection, we started with the base model and added in potential covariates one by one from our list of potential covariates that did not have significant missingness in order to improve the model fit. We using likelihood ratio tests at each step to determine if adding the covariate was an appropriate addition to the model (checking the reduced model against the full model).

### Exploratory analysis on covariates  

First we quantify the missingness in the outcomes:  

```{r}
tbl <- hrr %>% 
  select( `pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_40-64)`, `pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_65-74)`, `pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_75+)`, `pqi07_hypertension_admission_rate_(age_<_65)`,`pqi07_hypertension_admission_rate_(age_65-74)`,  `pqi07_hypertension_admission_rate_(age_75+)`, `pqi08_chf_admission_rate_(age_65-74)`, `pqi08_chf_admission_rate_(age_75+)`, `pqi08_chf_admission_rate_(age_<_65)`) %>% 
  summarize_all(.funs = function(x) return(paste0(round(sum(is.na(x))/length(x), 3)*100, "%"))) %>% t() %>% data.frame() 

names(tbl) <- c("% of Missing values")
kableExtra::kable(tbl)
```  

We have a very small percentage of missing data in our target covariates and 2 of the 3 PQIs. We believe PQI05 and PQI08 will be reasonable metrics to evaluate outpatient quality and it's relation to Medicaid expansion.

### Check distribution of outcomes  

#### PQI05  


```{r}
hrr %>% 
  filter(!is.na(expansion_status)) %>% 
  select(year, `pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_40-64)`, `pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_65-74)`, `pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_75+)`) %>% 
  gather(age, pqi05, -year) %>% 
  mutate(
    age = str_sub(age, -10)
  ) %>% ggplot(aes( x = pqi05)) + geom_histogram(color = "black") + 
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  xlab("PQI 05") +
  ylab("Count") +
  ggtitle("Distribution of PQI 05: COPD/Asthma in Older Adults Admissions Rate (2007-2018)") +
  theme_bw()
```    
The distribution has a slight right skewe.  


#### PQI08

```{r}
hrr %>% 
  filter(!is.na(expansion_status)) %>% 
  select(year, `pqi08_chf_admission_rate_(age_65-74)`, `pqi08_chf_admission_rate_(age_75+)`, `pqi08_chf_admission_rate_(age_<_65)`) %>% 
  gather(age, pqi08, -year) %>% 
  mutate(
    age = str_sub(age, -10)
  ) %>% ggplot(aes( x = pqi08)) + geom_histogram(color = "black") + 
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  xlab("PQI 08") +
  ylab("Count") +
  ggtitle("Distribution of PQI 05: Congestive Heart Failure Admissions Rate (2007-2018)") +
  theme_bw()

```    

This outcome is almost bi-modally normally distributed (or just has a large right skew).   



### Model Fitting  

#### PQI05

Since these variables are rates, we will want to use Poisson models. There may be overdispersion, so I'm going to start fitting a negative binomial and then check if the poisson version has overdispersion.  

```{r}
pqi05_data <- hrr %>% 
  filter(!is.na(expansion_status)) %>% 
  select(year, hrr, state, expansion_status,  percent_male, percent_african_american, `percent_hispanic`, percent_eligible_for_medicaid,`pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_40-64)`, `pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_65-74)`, `pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_75+)`) %>% 
  gather(age, pqi05, c("pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_40-64)", "pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_65-74)", "pqi05_copd_or_asthma_in_older_adults_admission_rate_(age_75+)")) %>% 
  mutate(age = str_extract(age, "(?<=age_).+(?=\\))"))
```     

##### Start with the base model  
```{r}
mod.nb_pqi05 <- MASS::glm.nb(pqi05 ~ `expansion_status` + offset(log(year)), data = pqi05_data)
summary(mod.nb_pqi05)
```  


We'll start adding in potential covariates and use likelihood ratio tests to determine if they should be added to the model.  

##### Percent Male  

```{r}
mod.nb_pqi05_c1 <- MASS::glm.nb(pqi05 ~ `expansion_status` +  percent_male + offset(log(year)), data = pqi05_data)
summary(mod.nb_pqi05_c1)

anova(mod.nb_pqi05, mod.nb_pqi05_c1)
```  
By the LRT, we add percent_male into the model.   

##### Percent African American  

```{r}
mod.nb_pqi05_c2 <- MASS::glm.nb(pqi05 ~ `expansion_status` +  percent_male + percent_african_american + offset(log(year)), data = pqi05_data)
summary(mod.nb_pqi05_c2)

anova(mod.nb_pqi05_c1, mod.nb_pqi05_c2)
```     
By the LRT, we add in percent African American into the model.  

##### Percent Hispanic  

```{r}
mod.nb_pqi05_c3 <- MASS::glm.nb(pqi05 ~ `expansion_status` +  percent_male + percent_african_american + percent_hispanic + offset(log(year)), data = pqi05_data)
summary(mod.nb_pqi05_c3)

anova(mod.nb_pqi05_c2, mod.nb_pqi05_c3)
```      
By the LRT, we add in percent hispanic into the model.  

##### Age category  

We add in the age category by which the data was originally split up.  

```{r}
mod.nb_pqi05_c4 <- MASS::glm.nb(pqi05 ~ `expansion_status` +  percent_male + percent_african_american + percent_hispanic + age + offset(log(year)), data = pqi05_data)
summary(mod.nb_pqi05_c4)

anova(mod.nb_pqi05_c3, mod.nb_pqi05_c4)
```      
By the LRT, we add age category into the model.  


##### Percent Eligible for Medicaid  
```{r}
mod.nb_pqi05_c5 <- MASS::glm.nb(pqi05 ~ `expansion_status` +  percent_male + percent_african_american + percent_hispanic + age + percent_eligible_for_medicaid + offset(log(year)), data = pqi05_data)
summary(mod.nb_pqi05_c5)

anova(mod.nb_pqi05_c4, mod.nb_pqi05_c5)
```       

By the LRT, we add in the percent eligible for Medicaid.  


##### Check for quadratic effects of `percent_eligible_for_medicaid`  

```{r}
mod.nb_pqi05_1 <- MASS::glm.nb(pqi05 ~ `expansion_status` +  percent_male + percent_african_american + percent_hispanic + age + percent_eligible_for_medicaid  + percent_eligible_for_medicaid_2+ offset(log(year)), data = pqi05_data %>% mutate(percent_eligible_for_medicaid_2 = percent_eligible_for_medicaid^2))
summary(mod.nb_pqi05_1)

anova(mod.nb_pqi05_c4, mod.nb_pqi05_1)
```    

By the LRT, we add the quadratic effects of the percent eligible for medicaid.   

#### Effect Modification  

We think there may be effect modification between % eligible for medicaid and expansion status, since in general Medicaid expansion means a higher percentage of the population qualifies for Medicaid.  

```{r}
mod.nb_pqi05_2 <-MASS::glm.nb(pqi05 ~ `expansion_status` +  percent_male + percent_african_american + percent_hispanic + age + percent_eligible_for_medicaid  + percent_eligible_for_medicaid_2 + expansion_status*percent_eligible_for_medicaid + offset(log(year)), data = pqi05_data %>% mutate(percent_eligible_for_medicaid_2 = percent_eligible_for_medicaid^2))
summary(mod.nb_pqi05_2 )  
anova(mod.nb_pqi05_1, mod.nb_pqi05_2)
```   

By the LRT, we do not add the interaction term to the model.  

#### Model Diagnostics  

```{r}
AIC(mod.nb_pqi05)
AIC(mod.nb_pqi05_c1)
AIC(mod.nb_pqi05_c2)
AIC(mod.nb_pqi05_c3)
AIC(mod.nb_pqi05_c4)
AIC(mod.nb_pqi05_c5)
AIC(mod.nb_pqi05_1)
```    

The full model has the lowest AIC.   

```{r}
BIC(mod.nb_pqi05)
BIC(mod.nb_pqi05_c1)
BIC(mod.nb_pqi05_c2)
BIC(mod.nb_pqi05_c3)
BIC(mod.nb_pqi05_c4)
BIC(mod.nb_pqi05_c5)
BIC(mod.nb_pqi05_1)
```     
The full model has the lowest AIC and BIC of all the models. 


##### Check Poisson model  

As a last step, we'll check the model as a Poisson model, and check if there is truly overdispersion.

```{r}
mod.pois_pqi05 <- glm(pqi05 ~ expansion_status + percent_male + 
    percent_african_american + percent_hispanic + age + percent_eligible_for_medicaid + 
    percent_eligible_for_medicaid_2 + 
    offset(log(year)), data = pqi05_data %>% mutate(percent_eligible_for_medicaid_2 = percent_eligible_for_medicaid^2), family = "poisson")

# Evaluate overdispersion
deviance(mod.pois_pqi05)/mod.pois_pqi05$df.residual #  78.17053
pearson.stat_pqi05 <- sum((na.omit(pqi05_data)$pqi05 - fitted(mod.pois_pqi05))^2/fitted(mod.pois_pqi05))
pearson.stat_pqi05/mod.pois_pqi05$df.residual # 79.35596

# Evidence of overdispersion 
mean(pqi05_data$pqi05, na.rm = T) # 866.4775
var(pqi05_data$pqi05, na.rm = T) # 157324.5
```  

By these, we can say the negative binomial is the right choice.   


#### Interpretation  

```{r}
exp(coef(mod.nb_pqi05_1))
exp(confint(mod.nb_pqi05_1))
```    


#### PQI08

Since these variables are rates, we will want to use Poisson models. There may be overdispersion, so I'm going to start fitting a negative binomial and then check if the poisson version has overdispersion.  

```{r}
pqi08_data <- hrr %>% 
  filter(!is.na(expansion_status)) %>% 
  select(year, hrr, state, expansion_status,  percent_male, percent_african_american, `percent_hispanic`, percent_eligible_for_medicaid,`pqi08_chf_admission_rate_(age_65-74)`, `pqi08_chf_admission_rate_(age_75+)`, `pqi08_chf_admission_rate_(age_<_65)`) %>% 
  gather(age, pqi08, c("pqi08_chf_admission_rate_(age_<_65)", "pqi08_chf_admission_rate_(age_65-74)", "pqi08_chf_admission_rate_(age_75+)")) %>% 
  mutate(age = str_extract(age, "(?<=age_).+(?=\\))"))
```     

##### Start with the base model  
```{r}
mod.nb_pqi08 <- MASS::glm.nb(pqi08 ~ `expansion_status` + offset(log(year)), data = pqi08_data)
summary(mod.nb_pqi08)
```  

We’ll start adding in potential covariates and use likelihood ratio tests to determine if they should be added to the model.

##### Percent Male   
```{r}
mod.nb_pqi08_c1 <- MASS::glm.nb(pqi08 ~ `expansion_status` +  percent_male + offset(log(year)), data = pqi08_data)
summary(mod.nb_pqi08_c1)

anova(mod.nb_pqi08, mod.nb_pqi08_c1)
```  
By the LRT, we add percent_male into the model.  

##### Percent African American   
```{r}
mod.nb_pqi08_c2 <- MASS::glm.nb(pqi08 ~ `expansion_status` +  percent_male + percent_african_american + offset(log(year)), data = pqi08_data)
summary(mod.nb_pqi08_c2)

anova(mod.nb_pqi08_c1, mod.nb_pqi08_c2)
```    


By the LRT, we add in percent African American into the model.   

##### Percent Hispanic

```{r}
mod.nb_pqi08_c3 <- MASS::glm.nb(pqi08 ~ `expansion_status` +  percent_male + percent_african_american + percent_hispanic + offset(log(year)), data = pqi08_data)
summary(mod.nb_pqi08_c3)  
anova(mod.nb_pqi08_c2, mod.nb_pqi08_c3)
```   

By the LRT, we add percent hispanic to the model.  

##### Age category  

```{r}
mod.nb_pqi08_c4 <- MASS::glm.nb(pqi08 ~ `expansion_status` +  percent_male + percent_african_american + percent_hispanic + age + offset(log(year)), data = pqi08_data)
summary(mod.nb_pqi08_c4) 
anova(mod.nb_pqi08_c3, mod.nb_pqi08_c4)
```  

By the LRT, we add in age category to the model.  

##### Percent Eligible for Medicaid  

```{r}
mod.nb_pqi08_c5 <- MASS::glm.nb(pqi08 ~ `expansion_status` +  percent_male + percent_african_american + percent_hispanic + age + percent_eligible_for_medicaid + offset(log(year)), data = pqi08_data)
summary(mod.nb_pqi08_c5)
anova(mod.nb_pqi08_c4, mod.nb_pqi08_c5)
```  

By the LRT we add in percent eligible for medicaid into the model.  

##### Check for quadratic effects of `percent_eligible_for_medicaid`   

```{r}
mod.nb_pqi08_1 <- MASS::glm.nb(pqi08 ~ `expansion_status` +  percent_male + percent_african_american + percent_hispanic + age + percent_eligible_for_medicaid  + percent_eligible_for_medicaid_2+ offset(log(year)), data = pqi08_data %>% mutate(percent_eligible_for_medicaid_2 = percent_eligible_for_medicaid^2))
summary(mod.nb_pqi08_1)  
anova(mod.nb_pqi08_c5, mod.nb_pqi08_1)
```    
By the LRT we add the quadratic effects of percent eligible for medicaid to the model.  

#### Effect Modification  

We think there may be effect modification between % eligible for medicaid and expansion status, since in general Medicaid expansion means a higher percentage of the population qualifies for Medicaid.  

```{r}
mod.nb_pqi08_2 <- MASS::glm.nb(pqi08 ~ `expansion_status` +  percent_male + percent_african_american + percent_hispanic + age + percent_eligible_for_medicaid  + percent_eligible_for_medicaid_2 + expansion_status*percent_eligible_for_medicaid + offset(log(year)), data = pqi08_data %>% mutate(percent_eligible_for_medicaid_2 = percent_eligible_for_medicaid^2))
summary(mod.nb_pqi08_2)  
anova(mod.nb_pqi08_1, mod.nb_pqi08_2)
```    

By the LRT, we don't add the interaction term to the model.  

#### Model Diagnostics  


```{r}
AIC(mod.nb_pqi08)
AIC(mod.nb_pqi08_c1)
AIC(mod.nb_pqi08_c2)
AIC(mod.nb_pqi08_c3)
AIC(mod.nb_pqi08_c4)
AIC(mod.nb_pqi08_c5)
AIC(mod.nb_pqi08_1)
```   

```{r}
BIC(mod.nb_pqi08)
BIC(mod.nb_pqi08_c1)
BIC(mod.nb_pqi08_c2)
BIC(mod.nb_pqi08_c3)
BIC(mod.nb_pqi08_c4)
BIC(mod.nb_pqi08_c5)
BIC(mod.nb_pqi08_1)
```  
The full model has the lowest AIC and BIC of all the models.

##### Check Poisson model  

```{r}
mod.pois_pqi08 <- glm(pqi08 ~ `expansion_status` + `percent_male` + `percent_african_american` + percent_hispanic + `percent_eligible_for_medicaid` + `age` + percent_eligible_for_medicaid_2, offset = log(year), data = pqi08_data %>% mutate(percent_eligible_for_medicaid_2 = percent_eligible_for_medicaid^2), family = "poisson")

# Evaluate overdispersion
deviance(mod.pois_pqi08)/mod.pois_pqi08$df.residual #  90.85094
pearson.stat_pqi08 <- sum((na.omit(pqi08_data)$pqi08 - fitted(mod.pois_pqi08))^2/fitted(mod.pois_pqi08))
pearson.stat_pqi08/mod.pois_pqi08$df.residual # 90.84562

# Evidence of overdispersion 
mean(pqi08_data$pqi08, na.rm = T) #  1447.86
var(pqi08_data$pqi08, na.rm = T) # 606848.1
```  

Due to the evidence of overdispersion, we can say the negative binomial is the right choice.   


#### Interpretation  

```{r}
exp(coef(mod.nb_pqi08_1))
exp(confint(mod.nb_pqi08_1))
```    


Create plots for report  

```{r}
pqi05_data %>% 
  na.omit() %>% 
  ggplot(aes(x = expansion_status, y = pqi05, fill = expansion_status)) +
  geom_violin(alpha = 0.5, draw_quantiles = c(0.5)) +
  xlab("Expansion Status") +
  ylab("COPD or asthma admissions in older\n adults per 1,000 Beneficiaries") + 
  scale_y_continuous(labels = scales::comma) +
  ggtitle("2018 Average PQI 05")
```  
```{r}
pqi08_data %>% 
  na.omit() %>% 
  ggplot(aes(x = expansion_status, y = pqi08, fill = expansion_status)) +
  geom_violin(alpha = 0.5, draw_quantiles = c(0.5)) +
  xlab("Expansion Status") +
  ylab("Congestive heart failure admissions per 1,000 beneficiaries") + 
  scale_y_continuous(labels = scales::comma) +
  ggtitle("2018 Average PQI 08")
```










