---
title: "Medicare Expansion Analysis: Part 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Data Prep

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(ggplot2)
library(gam)
library(corrplot)
```

```{r, echo=FALSE, include =FALSE}

# LOAD DATA ---------------------------------------------------------------


# Pull in 2018 demographic data for states
state_demo <- readxl::read_excel("./data/State County All Table 2018.xlsx", sheet = "State_county 2018", skip = 1)
# snake case field names
names(state_demo) <- tolower(str_replace_all(names(state_demo), " ", "_")) 


# CLEAN ROWS --------------------------------------------------------------

clean_nums_sing <- function(x) {
  tmp <- x
  # Clean percentages 
  if(!is.na(x)) {
    if(str_detect(x, "\\%")) {
      tmp <- str_replace(tmp, "\\%", "")
      tmp <- as.numeric(tmp)/100.0
      tmp <- as.character(tmp)
    }
    
    if(x == "*") {
      tmp <- str_replace_all(tmp, "\\*", "")
    }
  }
  return(tmp)
}

# function to modify columns
clean_nums <- function(col) {
return(sapply(col, clean_nums_sing))
}

state_demo <- state_demo %>% 
  mutate_all(clean_nums)


state_demo <- state_demo %>% 
  mutate_at(vars(`beneficiaries_with_part_a_and_part_b`:`pqi16_lower_extremity_amputation_admission_rate_(age_75+)`), as.numeric)


# SUBSET COLUMNS ----------------------------------------------------------
state_demo <- state_demo[, c("state", "county", "ffs_beneficiaries",  "ma_beneficiaries" , "ma_participation_rate", "average_age" , "percent_female", "percent_male" , "percent_non-hispanic_white", "percent_african_american"  , "percent_hispanic", "percent_other/unknown", "percent_eligible_for_medicaid", "average_hcc_score", "number_of_acute_hospital_readmissions" ,"hospital_readmission_rate", "emergency_department_visits", "emergency_department_visits_per_1000_beneficiaries", "%_of_beneficiaries_using_tests", "%_of_beneficiaries_using_imaging" , "%_of_beneficiaries_using_dme", "fqhc/rhc_visits_per_1000_beneficiaries", "%_of_beneficiaries_using_part_b_drugs")]



# ADD ON EXPANSION STATUS -------------------------------------------------

exp <- data.frame(state = state.abb, stringsAsFactors = F)

# KY, NV, CO, OR, NM, WV, AR, RI, AZ, MD, MA, ND, OH, IA, IL, VT, HI, NY, DE - 1/2014 Expansion states
exp_1_2014 <- c("KY", "NV", "CO", "OR", "NM", "WV", "AR", "RI", "AZ", "MD", "MA", "ND", "OH", "IA", "IL", "VT", "HI", "NY", "DE") 
# WA, CA, NJ, MN, DC, CT - Early expansion - (2010- 2013)
early_exp <- c("WA", "CA", "NJ", "MN", "DC", "CT" )
# NH, IN, MI, PA,AK,MT,LA - Late Expansion States (after Jan. 2014)
late_exp <- c("NH", "IN", "MI", "PA","AK","MT","LA" )

exp <- exp %>% 
  mutate(
    expansion_status = ifelse(state %in% early_exp, "1", ifelse(state %in% exp_1_2014, "2", ifelse(state %in% late_exp, "3", "4")))
  )


state_demo <- state_demo %>% 
  left_join(exp, by = "state")




```


```{r, echo=FALSE, include =FALSE}
state_demo %>% 
  filter(!county %in% c("NATIONAL TOTAL", "STATE TOTAL") & !is.na(expansion_status)) %>% 
  group_by(expansion_status) %>% 
  summarize(avg_mcd_elig = mean(percent_eligible_for_medicaid, na.rm = T),
            cnt_counties = n())

state_demo$exp <- ifelse(state_demo$expansion_status == "4", 0, 1)
state_demo$exp <- as.factor(state_demo$exp)
state_demo$not_exp <- ifelse(state_demo$expansion_status == "4", 1, 0)
state_demo$not_exp <- as.factor(state_demo$not_exp)

county_dat <- state_demo %>% filter(!county %in% c("NATIONAL TOTAL", "STATE TOTAL"))


county_dat <- county_dat %>% filter(!is.na(percent_eligible_for_medicaid) & 
                                    !is.na(emergency_department_visits_per_1000_beneficiaries) &
                                    !is.na(average_age) &
                                    !is.na(hospital_readmission_rate) & 
                                    !is.na(expansion_status))

```


## Linear Regression - HCC ~ expansion status

The hierarchical condition category (HCC) score is a model developed by 
CMS to assign risk scores for patients based on health conditions. The risk score
accounts for factors such as hospital admissions for acute and chronic issues, as well
as various classes of diseases and conditions ranging in severity. The
national average score is the reference point, and is equal to 1. A 68 year old individual
without any complications for their type-2 diabetes and a healthy BMI would be considered 
to have a total risk of 0, whereas a 68 year old patient with diabetic polyneuropathy, 
hypertension, obesity, and a below knee amputation would have a risk score of 1.1808
according to the CMS HCC model. 

Although it is commonly used by insurance companies to estimate future patient costs, this
metric can serve as a proxy to quantify the average individual health of a community. The average HCC score across U.S. counties is roughly normally distributed, with
a median value of 0.96. There distribution is slightly right skewed, and a log transform
might be considered if the model violates the LINE assumptions. 

### Exploratory Analysis
```{r}
summary(county_dat$average_hcc_score)
length(county_dat$average_hcc_score)
ggplot() + geom_histogram(aes(county_dat$average_hcc_score), col=1) 
           
```

```{r}
ggplot() + geom_histogram(aes(log10(county_dat$average_hcc_score)), fill=2) 
```

The primary covariate of interest is expansion status, and average age & percent 
eligible for medicaid are other important factors that should be adjusted for. There
seems to be a slight difference in average HCC scores in the states that have not expanded
& states that have expanded Medicaid, with larger variance in the counties that have 
not expanded. 
```{r}
county_dat  %>% ggplot() + geom_violin(aes(expansion_status, log10(average_hcc_score), 
                                          fill = expansion_status), alpha = 0.5) 
```

### Model Considerations

#### Assess Collinearity of Potential Covariates

```{r}

vars <- c("average_hcc_score",  
          "average_age",
          "percent_male",
          "percent_eligible_for_medicaid")
covars <- county_dat %>% select(vars) %>% drop_na()
corrplot(cor(covars), tl.cex = 0.7)

```



```{r}
mlr_red <- lm(average_hcc_score ~ as.factor(expansion_status), 
          data = county_dat)
summary(mlr_red)
```
```{r}
mlr <- lm(average_hcc_score ~ expansion_status +
                              percent_eligible_for_medicaid + 
                              I(percent_eligible_for_medicaid**2),
          data = county_dat)
summary(mlr)
plot(mlr)
```
### Model Diagnostics

The log transform does not remedy the normality violation of the model. 

```{r}
mlr_log <- lm(log(average_hcc_score) ~ expansion_status + 
                              # average_age + I(average_age**2) +
                              percent_eligible_for_medicaid + 
                              I(percent_eligible_for_medicaid**2),
          data = county_dat)
summary(mlr_log)
plot(mlr_log)
```

Consider average age & gender in model; both terms are significant &
improve Adjusted R squared value. However, the linearity assumption seems to
be violated based on the residuals vs fitted values plot. Also, the negative
coefficients seem to be counterintuitive (HCC score would be expected to increase
with increased age, not decrease as suggested by the negative coefficient).

```{r}
mlr2 <- lm(average_hcc_score ~ expansion_status + 
                              average_age + 
                              percent_male + 
                              percent_eligible_for_medicaid + 
                              I(percent_eligible_for_medicaid**2),
          data = county_dat)
summary(mlr2)
plot(mlr2)
```

The model with the additional linear & quadratic % eligible for medicaid
terms is needed to explain the variation in the data. The reduced model is not sufficient (F-test p < 0.05).

Additionally, the model with the age & gender terms explain more variation
in the data than the model consdering only the linear & quadratic % eligible for
medicaid terms.

```{r}
anova(mlr_red, mlr)
anova(mlr, mlr2)
```



```{r}
AIC(mlr2)
AIC(mlr)
```


#### Assess Effect Modification: 

The interaction terms are not statistically significant (p>0.05) when examining
whether expansion status modifies the % eligible for medicaid in each county. 


```{r}
mlr_em <- lm(average_hcc_score ~ expansion_status + 
                              average_age + I(average_age**2) +
                              percent_eligible_for_medicaid + 
                              expansion_status*percent_eligible_for_medicaid,
          data = county_dat)
summary(mlr_em)
```


The interaction term considering effect modification between average age & percent
male is significant, and also results in more intuitive coefficients that align
with expected directions for increasing males & older members of the county population.

The linearity of the model is also improved with the addition of the interaction term.

```{r}
mlr_em_age_gender <- lm(average_hcc_score ~ expansion_status + 
                              average_age +  #I(average_age**2) +
                              percent_male + I(percent_male**2) + 
                              percent_eligible_for_medicaid + 
                              I(percent_eligible_for_medicaid**2) + 
                              average_age*percent_male,
          data = county_dat)
summary(mlr_em_age_gender)
plot(mlr_em_age_gender)
```
```{r}
age <- 65
perc_male <- .5

# 1 unit change in % male
x <- 0.049072*age + 14.3*perc_male-7.57*perc_male**2 - 0.125855*age  

y <- 0.049072*age + 14.3*(perc_male + 0.01)-7.57*((perc_male + 0.01)**2) - 0.125855*age  

y - x

# 1 unit change in age
x <- 0.049072*age + 14.3*perc_male-7.57*perc_male**2 - 0.125855*age  

y <- 0.049072*(age+1) + 14.3*perc_male-7.57*(perc_male**2) - 0.125855*(age+1)  

y - x
```

```{r}
m <- lm(average_hcc_score ~ percent_male + I(percent_male**2) + 
                            average_age + 
                            percent_male*average_age,
           data = county_dat)

plot(county_dat$percent_male, county_dat$average_hcc_score, col = 8, 
     xlab = "percent male", ylab = "average hcc score")
lines(seq(.3,.7,.05), predict(m, newdata = data.frame(average_age = 65, 
                                                       percent_male = seq(.3,.7,.05))),
      col = 3
      )

lines(seq(.3,.7,.05), predict(m, newdata = data.frame(average_age = 71, 
                                                       percent_male = seq(.3,.7,.05))),
      col = 5
      )

lines(seq(.3,.7,.05), predict(m, newdata = data.frame(average_age = 72, 
                                                       percent_male = seq(.3,.7,.05))),
      col = 4
      )
legend("topleft", c("65 years", "71 years", "72 years"), pch=1, col=c(3,5,4))

```


```{r}
AIC(mlr_em_age_gender)
```

```{r}
anova( mlr2, mlr_em_age_gender)
```

#### Robust Sandwich Errors for Coefficient Estimates

Because the residual plot depicts a violation of homoscedasticity and the qq plot
illustrates deviation from normality, robust sandwich estimation methods were used to
calculate the standard errors for the coefficients in the
final adjusted model. This method is more robust than the
standard LSE method for estimating the standard errors, and can be used to derive more
accurate confidence intervals. 

```{r}
library(lmtest)
library(sandwich)
coeftest(mlr_em_age_gender, vcov = vcovHC(mlr_em_age_gender, type="HC1"))

```

```{r}
# Confidence Intervals
(0.0010648  + c(-1,1)*1.96*0.0052007) #95% ci for expansion status 2
(0.0145870 + c(-1,1)*1.96*0.0057930 ) #95% ci for expansion status 3
(0.0345069 + c(-1,1)*1.96* 0.0051904) #95% ci for expansion status 4



```


### D. Model Summary 

$$
\begin{split}
& E(\text{average HCC score}) = -3.54 \\
& 0.001*\text{I(State Expanded Medicaid in 2014)} + \\
& 0.015*\text{I(State Expanded Medicaid after 2014)} + \\
& 0.035*\text{I(State has not Expanded Medicaid)} + \\
& 0.049*\text{(average age)} + \\
& 14.31*\text{(percent male)}  \\
& -7.58*\text{(percent male)}^2 + \\
& 0.82*\text{(percent eligible for medicaid)}  \\
& -0.45*\text{(percent eligible for medicaid)}^2  \\
& -0.13*\text{(average age * percent male)} 
\end{split}
$$ 


## Linear Regression: - HCC 2018 vs 2010

Is there a difference in average HCC scores in 2018 vs 2010 for counties
that expanded Medicaid early versus late?

```{r, echo = FALSE, include = FALSE}

# LOAD DATA ---------------------------------------------------------------


# Pull in 2018 demographic data for states
state_demo_2010 <- readxl::read_excel("./data/State County All Table 2018.xlsx", sheet = "State_county 2010", skip = 1)
# snake case field names
names(state_demo_2010) <- tolower(str_replace_all(names(state_demo_2010), " ", "_")) 


# CLEAN ROWS --------------------------------------------------------------

state_demo_2010 <- state_demo_2010 %>% 
  mutate_all(clean_nums)

state_demo_2010 <- state_demo_2010 %>% 
  mutate_at(vars(`beneficiaries_with_part_a_and_part_b`:`pqi16_lower_extremity_amputation_admission_rate_(age_75+)`), as.numeric)


# SUBSET COLUMNS ----------------------------------------------------------
state_demo_2010 <- state_demo_2010[, c("state", "county", 
                                       "average_age" , "percent_female", "percent_male" ,
                                       "percent_eligible_for_medicaid", "average_hcc_score")]


```

```{r, echo = FALSE, include = FALSE}
state_demo_combo <- state_demo %>%
  inner_join(state_demo_2010, by = c("state", "county"), 
            suffix = c("_2018", "_2010"))


state_demo_combo$exp <- ifelse(state_demo_combo$expansion_status == "4", 0, 1)
state_demo_combo$exp <- as.factor(state_demo_combo$exp)
state_demo_combo$not_exp <- ifelse(state_demo_combo$expansion_status == "4", 1, 0)
state_demo_combo$not_exp <- as.factor(state_demo_combo$not_exp)

county_dat_combo <- state_demo_combo %>% filter(!county %in% c("NATIONAL TOTAL", "STATE TOTAL"))


county_dat_combo['hcc_ratio'] = county_dat_combo$average_hcc_score_2018/county_dat_combo$average_hcc_score_2010

county_dat_combo['hcc_diff'] = county_dat_combo$average_hcc_score_2018 -county_dat_combo$average_hcc_score_2010


county_dat_combo <- county_dat_combo %>% mutate(
            delta_male = (percent_male_2018 - percent_male_2010)/percent_male_2010, 
            delta_age = (average_age_2018 - average_age_2010)/average_age_2010,
            delta_medicaid = (percent_eligible_for_medicaid_2018 - percent_eligible_for_medicaid_2010)/percent_eligible_for_medicaid_2010)


county_dat_combo <- county_dat_combo %>% filter(!is.na(expansion_status) &
                                                !is.na(delta_age) & 
                                                !is.na(delta_medicaid) &
                                                !is.na(delta_male))
```


### Exploratory Analysis

Both the hcc ratio of 2018 vs 2010 values and the difference are normally distributed with long tails on both ends. Taking the log transform of the ratio
does not substantially remedy the long tail issue.

```{r}


summary(county_dat_combo$hcc_ratio)
ggplot() + geom_histogram(aes(county_dat_combo$hcc_ratio), col=1)

summary(county_dat_combo$hcc_ratio)
ggplot() + geom_histogram(aes(log10(county_dat_combo$hcc_ratio)), col=1)

summary(county_dat_combo$hcc_diff)
ggplot() + geom_histogram(aes(county_dat_combo$hcc_diff), col=1)
```


```{r}
comp_red <- lm(hcc_diff ~ as.factor(expansion_status), 
          data = county_dat_combo)
summary(comp_red)
```
```{r}
comp1 <- lm(hcc_diff ~ as.factor(expansion_status) +
                       delta_medicaid + 
                       delta_age +
                       delta_male , 
          data = county_dat_combo)
summary(comp1)
plot(comp1)
```
The additional quadratic delta_medicaid term improves the linearity of the final
model, and is statistically significant.

```{r}
comp2 <- lm(hcc_diff ~ as.factor(expansion_status) +
                       delta_medicaid + 
                       delta_age +
                       delta_male + 
                       I(delta_medicaid**2) , 
          data = county_dat_combo)
summary(comp2)
plot(comp2)
```

The full model with the additional delta age, percent male, and percent eligible
for medicaid is needed to explain the variation in the data.

```{r}
anova(comp_red, comp2)
```
#### Robust Sandwich Errors for Coefficient Estimates

Because the residual plot depicts a violation of homoscedasticity and the qq plot
illustrates deviation from normality, robust sandwich estimation methods were used to
calculate the standard errors for the coefficients in the
final adjusted model. This method is more robust than the
standard LSE method for estimating the standard errors, and can be used to derive more
accurate confidence intervals. 

```{r}
library(lmtest)
library(sandwich)
coeftest(comp2, vcov = vcovHC(comp2, type="HC1"))

```
```{r}
# Confidence Intervals
(0.0196943  + c(-1,1)*1.96*0.0052007) #95% ci for expanded vs not




```
## Logistic Regression

Create binary outcome categorizing hospital readmission rates & emergency department visits as high if the county rate exceeds the national average rate (hospital readmission - 18.06%, ED visits - 670 per 1000 beneficiaries)

Also, create multinomial class var with following labels:

     * 4 - hospital readmission rate & % beneficiaries with an ED visit above national average
     * 3 - beneficiaries w/ ED visit above national average (hospital readmission below)
     * 2 - hospital readmission rate above national average (ED below)
     * 1  - both hospital readmission & ED below
    
```{r}

county_dat['high_hospital_readmission'] = ifelse(county_dat$hospital_readmission_rate >= .1806, 1, 0)

county_dat['high_ED'] = ifelse(county_dat$emergency_department_visits_per_1000_beneficiaries >= 670, 1, 0)

county_dat['hospital_ED_class'] = case_when(
                                    (county_dat$high_hospital_readmission == 1) &
                                    (county_dat$high_ED == 1)~ 3,
                                    county_dat$high_ED == 1 ~ 2,
                                    county_dat$high_hospital_readmission == 1 ~ 2,
                                    TRUE ~ 1)


```


#### Emergency Department Visits

Expansion status of state is only predictive in the logistic model 
looking at high ED rates in counties; it is not a statistically significant 
predictor in the model with high hospitalization rates as the outcome (neither 
in binary - expanded vs non-expanded nor more verbose 4/3/2/1 classification)

```{r}
mod.binary.ed = glm(high_ED ~ 
                 as.factor(expansion_status) + 
                 percent_male +  
                 average_age + 
                 percent_eligible_for_medicaid +
                I(percent_eligible_for_medicaid**2),
                 family = binomial(), data = county_dat)
summary(mod.binary.ed)
```

```{r}
mod.binary.ed.cont = glm(high_ED ~ 
                 as.integer(expansion_status) + 
                 percent_male +  
                 average_age + 
                 percent_eligible_for_medicaid +
                 I(percent_eligible_for_medicaid**2),
                 family = binomial(), data = county_dat)
summary(mod.binary.ed.cont)

```
Reduced model treating expansion_status as a continous covariate is sufficient;
the categorical information is not needed.

```{r}
anova(mod.binary.ed.cont, mod.binary.ed, test="Chisq")
```


```{r}
exp(0.44094 + c(-1,1)*0.18333*1.96) # expansion status 2 vs 1
exp(0.37138 + c(-1,1)*0.20519*1.96) # expansion status 3 vs 1
exp(0.52777 + c(-1,1)*0.17554*1.96) # expansion status 4 vs 1


# continuous
exp(0.10027 + c(-1,1)*0.04257*1.96)

```
# GOF - Hosmer Lemeshow Goodness of Fit
```{r}
library(ResourceSelection)

hoslem.test(county_dat$high_ED, predict(mod.binary.ed), g = 10)
hoslem.test(county_dat$high_ED, predict(mod.binary.ed.cont), g= 10)
```

```{r}
library(pROC)

roc_cont <- roc(county_dat$high_ED, predict(mod.binary.ed.cont))
roc_cat <- roc(county_dat$high_ED, predict(mod.binary.ed))
plot(roc_cont)
plot(roc_cat, add =TRUE, col = 3)

legend(0,0.5, legend=c("continuous", "categorical"), col=c(1,3), lty=1)

auc(roc_cont)
auc(roc_cat)

```


#### Hospital Readmissions
```{r}
mod.binary.hosp = glm(high_hospital_readmission ~ 
                 as.factor(expansion_status) + 
                 percent_male +  
                 average_age + 
                 percent_eligible_for_medicaid,
                 family = binomial(), data = county_dat)
summary(mod.binary.hosp)
```


## Multinomial Regression

```{r}
library(nnet)
mod.multi <- multinom(as.factor(hospital_ED_class) ~ 
                 as.factor(expansion_status) + 
                 percent_male +  
                 average_age + 
                 percent_eligible_for_medicaid +
                 I(percent_eligible_for_medicaid**2) +
                 I(average_age**2),
                 data = county_dat)

summary(mod.multi)
```

The model is not doing a good job classifying health outcome class
based on expansion status.

```{r}
county_dat['predict'] <- predict(mod.multi)
county_dat %>% ggplot() + geom_histogram(aes(predict, fill = expansion_status), stat="count")

```

```{r}
county_dat %>% ggplot() + geom_histogram(aes(hospital_ED_class, fill = expansion_status), stat="count")
```

## Ordinal Regression
```{r}
library(VGAM)
ord.all <- vglm(hospital_ED_class ~ 
                 expansion_status + 
                 percent_male +  
                 average_age + 
                 percent_eligible_for_medicaid +
                 I(percent_eligible_for_medicaid**2), 
                cumulative(parallel=TRUE, reverse=TRUE), 
                data=county_dat)

summary(ord.all)

```

Test proportional odds assumption - no overlap in beta estimates confidence interval. Proportional odds assumption is violated.
```{r}
county_dat['ind3'] = ifelse(county_dat$hospital_ED_class == 3, 1, 0)
county_dat['ind32'] = ifelse(county_dat$hospital_ED_class == 1, 0, 1)

mod.3v12 <- glm(ind3 ~expansion_status + 
                 percent_male +  
                 average_age + 
                 percent_eligible_for_medicaid +
                 I(percent_eligible_for_medicaid**2),
                 family=binomial,
                 data=county_dat)
                
                

mod.32v1 <- glm(ind32 ~expansion_status + 
                 percent_male +  
                 average_age + 
                 percent_eligible_for_medicaid +
                 I(percent_eligible_for_medicaid**2),
                 family=binomial,
                 data=county_dat)

summary(mod.3v12)
summary(mod.32v1)
                
```


```{r}

ord.all_potest <- vglm(hospital_ED_class ~ 
                 as.factor(expansion_status) + 
                 percent_male +  
                 average_age + 
                 percent_eligible_for_medicaid +
                 I(percent_eligible_for_medicaid**2), 
                cumulative(parallel=FALSE, reverse=T), 
                data=county_dat)

summary(ord.all_potest)
```

