---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Prep

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(ggplot2)

```


```{r, echo=FALSE, include =FALSE}

# LOAD DATA ---------------------------------------------------------------


# Pull in 2018 demographic data for states
state_demo <- readxl::read_excel("./data/State County All Table 2018.xlsx", sheet = "State_county 2018", skip = 1)
# snake case field names
names(state_demo) <- tolower(str_replace_all(names(state_demo), " ", "_")) 


# CLEAN ROWS --------------------------------------------------------------

clean_nums_sing <- function(x) {
  tmp <- x
  # Clean percentages 
  if(!is.na(x)) {
    if(str_detect(x, "\\%")) {
      tmp <- str_replace(tmp, "\\%", "")
      tmp <- as.numeric(tmp)/100.0
      tmp <- as.character(tmp)
    }
    
    if(x == "*") {
      tmp <- str_replace_all(tmp, "\\*", "")
    }
  }
  return(tmp)
}

# function to modify columns
clean_nums <- function(col) {
return(sapply(col, clean_nums_sing))
}

state_demo <- state_demo %>% 
  mutate_all(clean_nums)


state_demo <- state_demo %>% 
  mutate_at(vars(`beneficiaries_with_part_a_and_part_b`:`pqi16_lower_extremity_amputation_admission_rate_(age_75+)`), as.numeric)


# SUBSET COLUMNS ----------------------------------------------------------
state_demo <- state_demo[, c("state", "county", "ffs_beneficiaries",  "ma_beneficiaries" , "ma_participation_rate", "average_age" , "percent_female", "percent_male" , "percent_non-hispanic_white", "percent_african_american"  , "percent_hispanic", "percent_other/unknown", "percent_eligible_for_medicaid", "number_of_acute_hospital_readmissions" ,"hospital_readmission_rate", "emergency_department_visits", "emergency_department_visits_per_1000_beneficiaries", "%_of_beneficiaries_using_tests", "%_of_beneficiaries_using_imaging" , "%_of_beneficiaries_using_dme", "fqhc/rhc_visits_per_1000_beneficiaries", "%_of_beneficiaries_using_part_b_drugs", "pqi03_diabetes_lt_complication_admission_rate_(age_<_65)")]



# ADD ON EXPANSION STATUS -------------------------------------------------

exp <- data.frame(state = state.abb, stringsAsFactors = F)

# KY, NV, CO, OR, NM, WV, AR, RI, AZ, MD, MA, ND, OH, IA, IL, VT, HI, NY, DE - 1/2014 Expansion states
exp_1_2014 <- c("KY", "NV", "CO", "OR", "NM", "WV", "AR", "RI", "AZ", "MD", "MA", "ND", "OH", "IA", "IL", "VT", "HI", "NY", "DE") 
# WA, CA, NJ, MN, DC, CT - Early expansion - (2010- 2013)
early_exp <- c("WA", "CA", "NJ", "MN", "DC", "CT" )
# NH, IN, MI, PA,AK,MT,LA - Late Expansion States (after Jan. 2014)
late_exp <- c("NH", "IN", "MI", "PA","AK","MT","LA" )

exp <- exp %>% 
  mutate(
    expansion_status = ifelse(state %in% early_exp, "1", ifelse(state %in% exp_1_2014, "2", ifelse(state %in% late_exp, "3", "4")))
  )


state_demo <- state_demo %>% 
  left_join(exp, by = "state")




```

```{r}
state_demo %>% 
  filter(!county %in% c("NATIONAL TOTAL", "STATE TOTAL") & !is.na(expansion_status)) %>% 
  group_by(expansion_status) %>% 
  summarize(avg_mcd_elig = mean(percent_eligible_for_medicaid, na.rm = T),
            cnt_counties = n())

state_demo$exp <- ifelse(state_demo$expansion_status == "4", 0, 1)
state_demo$exp <- as.factor(state_demo$exp)
state_demo$not_exp <- ifelse(state_demo$expansion_status == "4", 1, 0)
state_demo$not_exp <- as.factor(state_demo$not_exp)

county_dat <- state_demo %>% filter(!county %in% c("NATIONAL TOTAL", "STATE TOTAL"))


county_dat <- county_dat %>% filter(!is.na(percent_eligible_for_medicaid) & 
                                    !is.na(emergency_department_visits_per_1000_beneficiaries) &
                                    !is.na(average_age) &
                                    !is.na(hospital_readmission_rate))

```


## Data Analysis

```{r, echo=FALSE, include=FALSE}
names(county_dat)
```

#### Binomial Logistic Regression


Initially we planned to examine Preventive Quality Indicators as a measure of health quality
in a population, but these data points are only publically available at the state level, which
would limit our analysis to only 50 observations. Thus, we have modified our analysis approach to
consider data at the county level and use alternative measures for quality of health.
```{r}
pqi3 <- state_demo %>% filter(county == 'STATE TOTAL') %>% rename(pqi_03_less_65 = 'pqi03_diabetes_lt_complication_admission_rate_(age_<_65)')

summary(glm(not_exp ~ pqi_03_less_65, family = binomial(), data=pqi3))

```


This reduced model confirms that the precentage of people eligible for medicaid in the Medicare
population (i.e. dual eligible) is associated with whether the state has expanded 
Medicaid access as part of the Affordable Care Act. 

```{r}
mod.1 <- glm(not_exp ~ percent_eligible_for_medicaid, family = binomial(), 
            data= county_dat)
summary(mod.1)
```


This extended model considers the additional covariates of health quality measures
(hospital readmission rate & number of emergency department visits) and also controls for average
age in the county population when predicting the county's Medicaid expansion status.

```{r}
mod.binary = glm(not_exp ~ 
                 emergency_department_visits_per_1000_beneficiaries +
                 hospital_readmission_rate + 
                 average_age + 
                 percent_eligible_for_medicaid,
                 family = binomial(), data = county_dat)
summary(mod.binary)

```

```{r}
exp(100*0.0015904)
CI=c(exp(100*(0.0015904-1.96*0.0003563)),exp(100*(0.0015904+1.96*0.0003563)))
CI


exp(2.6013322)
CI=c(exp((2.6013322-1.64*1.4463722)),exp((2.6013322+1.64*1.4463722)))
CI


```


```{r}
anova(mod.1, mod.binary, test="Chisq" )

```


#### Multinomial  Logistic Regression (y - expansion status category of 1, 2, 3, or 4)
```{r}

library(nnet)
mod.all_multi <- multinom(expansion_status ~ hospital_readmission_rate +
                          average_age  +  
                          emergency_department_visits_per_1000_beneficiaries + 
                          percent_eligible_for_medicaid, 
                          data = county_dat)
summary(mod.all_multi)
```
```{r}
exp(2.49)
CI=c(exp((2.49-1.96*6.501529e-06)),exp((2.49+1.96*6.501529e-06)))
CI


exp(-0.57)
CI=c(exp((-0.57-1.96* 6.783909e-06)),exp((-0.57+1.96* 6.783909e-06)))
CI

exp(3.95)
CI=c(exp((3.95-1.96*5.847374e-06)),exp((3.95+1.96*5.847374e-06)))
CI



```


#### Ordinal Logistic Regression

```{r}
library(VGAM)

mod.no.po <- vglm(as.numeric(expansion_status) ~ emergency_department_visits_per_1000_beneficiaries +
                  average_age ,
                  cumulative(parallel=FALSE, reverse=T), data=county_dat)

summary(mod.no.po)

```

