---
title: "Potential Confounders"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
medicare <- read_csv("data/medicare_2007_2018.csv")
medicare$expansion_status <- factor(medicare$expansion_status)

hrr <- read_csv("data/hrr_all_years.csv")
hrr$expansion_status <- factor(hrr$expansion_status)
```   
## GVPUF - County

### Potential Covariates  
We evaluated what percentage of each of our covariates of interest were missing:  

```{r}
tbl <- medicare %>% 
  select(
    county, average_age, percent_male, `percent_non-hispanic_white`, 
    percent_african_american, percent_hispanic, 
    percent_eligible_for_medicaid
  ) %>% 
  summarize_all(.funs = function(x) 
    return(paste0(round(sum(is.na(x))/length(x), 3)*100, "%"))) %>% 
  t() %>% data.frame() 

names(tbl) <- c("% of Missing values")
kableExtra::kable(tbl)
```    

We investigated the pattern of missingness:  

```{r}
medicare %>% 
  mutate(
    p_white_missing = is.na(`percent_non-hispanic_white`),
    p_aa_missing = is.na(percent_african_american),
    p_h_missing = is.na(percent_hispanic)
  ) %>% 
  group_by(year, p_white_missing, p_aa_missing, p_h_missing) %>% 
  summarize(cnt = n())
```   

It appears that the same counties do not report these metrics. Given the high percentage of missingness in these covariates, I've decided to drop them from my potential covariates list.  

#### Check correlaton in the covariates  

Complete cases with race variables

```{r}
medicare %>% 
  select(
    average_age, percent_male, `percent_non-hispanic_white`, 
    percent_african_american, percent_hispanic, 
    percent_eligible_for_medicaid
  ) %>% na.omit() %>% 
  cor() %>% corrplot::corrplot(method = "number")
```     
There is a strong correlation between percent African American and percent non-Hispanic white.  

Complete cases without race variables

```{r}
medicare %>% 
  select(
    average_age, percent_male, percent_eligible_for_medicaid
  ) %>% na.omit() %>% 
  cor() %>% 
  corrplot::corrplot(method = "number")
```      
Without the race variables, which have a high amount of missingness, there is a high correlation between percent eligible for Medicaid and average age.  

## GVPUF - HRR

### Potential Covariates  
We evaluated what percentage of each of our covariates of interest were missing:  

```{r}
tbl <- hrr %>% 
  select(
    hrr, average_age, percent_male, `percent_non-hispanic_white`, percent_african_american, percent_hispanic, percent_eligible_for_medicaid
  ) %>% 
  summarize_all(.funs = function(x) 
    return(paste0(round(sum(is.na(x))/length(x), 3)*100, "%"))) %>% 
  t() %>% data.frame() 

names(tbl) <- c("% of Missing values")
kableExtra::kable(tbl)
```    

There is a very small amount of missingness across this data.  

#### Check correlaton in the covariates  

```{r}
hrr %>% 
  select(
    average_age, percent_male, `percent_non-hispanic_white`, 
    percent_african_american, percent_hispanic, 
    percent_eligible_for_medicaid
  ) %>% na.omit() %>% 
  cor() %>% corrplot::corrplot(method = "number")
```     
There is a high amount of correlation between percent Hispanic and percnet non-Hispanic white. 

